Predict Models

### 2024-01-06 scripts/predict_models.Rmd
pushed processed onto ssd, avoiding parallel to max RAM on single thread

```{r}
library(tidyverse)
library(terra)
library(sf)
source('functions.R')
```

```{r read in predictors}

setGDALconfig(c("GDAL_MAX_DATASET_POOL_SIZE=1024"))

p <- '/media/steppe/hdd/Geospatial_data/WesternPlantPredictors2'
paths <- list.dirs(p, recursive = T)
paths <- paths[2:length(paths)]
paths <- paths[! grepl('chelsa$|veg_cover$|soilgrids$|sand$|silt$|clay$|phh2o$|cfvo$|soc$', paths) ]

@@

pmodels <- '../results/rf_models'
f <- file.path(pmodels, list.files(pmodels))

@@

  pout <- '../results/suitability_maps'
  layer_subset <- terra::subset(layers, term_grab(sdm))
  layer_subset <- terra::crop(layer_subset, bound)

  return(layer_subset)
  #terra::predict(
  #  type = 'prob', cores = 16,
  #  layer_subset, sdm, cpkgs="randomForest",
  #  filename = file.path(pout, paste0(x$species, '.tif')),
  #  wopt = c(names = 'predicted_suitability'))

@@

ob <- lapply(models_occ[c(5)], predict_wrapper)
ob1 <- ob[[1]]
ob1

predict_wrapper(models_occ[[5]])
```

@@

>>>>

```{r}
library(tidyverse)
library(terra)
library(sf)
library(randomForest)
source('functions.R')

# unix::rlimit_as(1e32)
# setGDALconfig(c("GDAL_MAX_DATASET_POOL_SIZE"), c(""))
```

```{r read in predictors}
p <- '~/Documents/WesternPlantPredictors2'
paths <- list.dirs(p, recursive = T)
paths <- paths[2:length(paths)]
paths <- paths[! grepl('chelsa$|veg_cover$|soilgrids$|sand$|silt$|clay$|phh2o$|cfvo$|soc$', paths) ]

@@

pmodels <- '../results/rf_models'
f <- file.path(pmodels, list.files(pmodels))
f <- data.frame(
  path = f,  
  species = gsub('-.*$', '', basename(f)), 
  time = gsub('[A-Za-z]', '', basename(f))
  ) |>
  mutate(time = gsub('_-|[.]', '', time),
         time = gsub('_', ' ', time), 
         time = as.POSIXct(time)) |>
  group_by(species) |>
  slice_max(order_by = time, n = 1) |>
  pull(path)

@@

  pout <- '~/Documents/suitability_maps'
  layer_subset <- terra::subset(layers, term_grab(sdm))
  layer_subset <- terra::crop(layer_subset, bound)

  OneClass <- \(...) predict(...)[,2]
  terra::predict(
    layer_subset, sdm, cpkgs="randomForest",
    type = 'prob', cores = 1, f = OneClass,
    filename = file.path(pout, paste0(x$species, '-', gsub(' ', '_', Sys.time()), '.tif')),
    wopt = c(names = 'predicted_suitability'))

@@

lapply(models_occ[2], predict_wrapper)

```

@@

```{r}
logo <- rast(system.file("ex/logo.tif", package="terra"))   
names(logo) <- c("red", "green", "blue")
p <- matrix(c(48, 48, 48, 53, 50, 46, 54, 70, 84, 85, 74, 84, 95, 85, 
              66, 42, 26, 4, 19, 17, 7, 14, 26, 29, 39, 45, 51, 56, 46, 38, 31, 
              22, 34, 60, 70, 73, 63, 46, 43, 28), ncol=2)

a <- matrix(c(22, 33, 64, 85, 92, 94, 59, 27, 30, 64, 60, 33, 31, 9,
              99, 67, 15, 5, 4, 30, 8, 37, 42, 27, 19, 69, 60, 73, 3, 5, 21,
              37, 52, 70, 74, 9, 13, 4, 17, 47), ncol=2)

xy <- rbind(cbind(1, p), cbind(0, a))

# extract predictor values for points
e <- extract(logo, xy[,2:3])

# combine with response (excluding the ID column)
v <- data.frame(cbind(pa=xy[,1], e))

rm(p, a, xy, e)

### with two output variables (probabilities for each class)
v$pa <- as.factor(v$pa)
rfm2 <- randomForest(formula=pa~., data=v)
rfp <- predict(logo, rfm2, cores=2, type="prob", cpkgs="randomForest")

f <- \(...) predict(...)[,2]
rfp <- predict(logo, rfm2, cores=2, type="prob", cpkgs="randomForest", fun=f)

```
